---
title: "Attempted fits to Hypotheses"
author: "Pi Nuessle and Judith Racusin"
date: '2023-06-09'
output: word_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(plotly)
library(tidyr)
library("scatterplot3d")
# library(randomForest)
library(e1071)
library(misc3d)
library(plot3D)
library(scatterplot3d)
library(rgl)
library(data.table)
library(caret)
# library(tidyverse)
set.seed(1729)
```

```{r defining the special bursts, echo=FALSE}
Collapsars = c(050416461.0, 50525002.0, 050824966.0, 060218148.0, 060729800.0, 060904104.0, 070419447.0, 071025172.0, 071112772.0, 080319258.0, 081007224.0, 090618353.0, 091127976.0, 100316531.0, 100418882.0, 101219686.0, 101225776.0, 111209300.0, 111211928.0, 111228656.0, 120422300.0, 120714888.0, 120729455.0, 130215063.0, 130427324.0, 130702003.0, 130831544.0, 140206303.0, 140606133.0, 150818483.0, 161219783.0, 161228552.0, 171010792.0, 171205306.0, 180720598.0, 180728728.0, 190114872.0, 190829830.0, 200826187.0, 210210083.0, 211023545.0, 221009553.0, 200826187.0)
Mergers=c(070809807.0, 130603659.0, 150101641.0, 160624477.0, 160821936.0, 170817528.0, 200522487.0, 060614530.0, 111005336.0, 120304248.0, 211211549.0, 230307655.0, 050509166.0, 050709942.0, 051210240.0, 070714207.0, 071227842.0, 080503518.0, 080905499.0, 090515198.0, 160303454.0)
```


The chunk that follows does a linear fit of the prompt isotropic energy against the average decay rate (possibly with redshift). While it is successful, the fit may not be significantly different from fitting nothing. I can actually prove this.

```{r Hypothesis 1, echo=FALSE, eval=FALSE}
H_1_Data <- read_excel("/Users/nnuessle/Documents/Work/burst_matching_algorithm/actual_paper/prompt-and-afterglow-matching/Hypothesis_1_Data.xlsx", col_types = c("skip", "numeric", "numeric", "numeric", "numeric", "numeric"))
H_1_Data_copy<-data.frame(H_1_Data)
#names(H_1_Data_copy) <- NULL             # Delete column names
#H_1_Data_copy                            # Print updated data
# cat("we had to remove bursts ")
# # cat(toString(H_1_Data[16, 'Name']))
# # cat(', ')
# cat(toString(H_1_Data[21, 'Name']))
# cat(', and')
# cat(toString(H_1_Data[46, 'Name']))
# cat(", because they were outliers in the linear model.")
# H_1_Data_copy<-H_1_Data_copy[-c(10, 19, 27, 40, 43), ]
H_1_Data_copy<-H_1_Data_copy[-c(16), ]
H_1_Data_copy$Prompt.Isotropic.Energy<-log10(H_1_Data_copy$Prompt.Isotropic.Energy)
x<-H_1_Data_copy$Prompt.Isotropic.Energy
y<-H_1_Data_copy$Avg..Afterglow.Decay.Rate
z<-H_1_Data_copy$Redshift
mod1<-lm(y~x)
#mod1<-lm(y~x+z)
par(mfrow=c(2,2))
plot(mod1)
anova(mod1)
print(mod1)
newx <- seq(min(x), max(x), length.out=100)
# newz <- seq(min(z), max(z), length.out=100)
preds <- predict(mod1, newdata=data.frame(x=newx), interval='prediction')
# preds <- predict(mod1, newdata=data.frame(x=newx, z=newz), interval='confidence')
par(mar=c(5,5,5,5))
par(mfrow=c(1,1))
collapsar_mask=H_1_Data_copy$Name %in% Collapsars
merger_mask=H_1_Data_copy$Name %in% Mergers
H_1_Data_collapsars=H_1_Data_copy[collapsar_mask, ]
H_1_Data_mergers=H_1_Data_copy[merger_mask, ]
H_1_Data_unknown=H_1_Data_copy[!c(collapsar_mask+merger_mask), ]
plot(H_1_Data_unknown$Avg..Afterglow.Decay.Rate~H_1_Data_unknown$Prompt.Isotropic.Energy, xlab='Prompt Isotropic Energy 10kev-10MeV (erg, GBM)', ylab=paste(strwrap('Redshift-Corrected Afterglow Decay Rate 0.2-10 keV (photons s^-2, XRT)', width = 50), collapse = "\n"), main=paste(strwrap('Isotropic-Equivalent Energy and Average Afterglow Decay Lack of a Relationship', width = 50), collapse = "\n"), pch=20)
# points(H_1_Data_mergers$Avg..Afterglow.Decay.Rate~H_1_Data_mergers$Prompt.Isotropic.Energy, pch=4, col='red')
points(H_1_Data_collapsars$Avg..Afterglow.Decay.Rate~H_1_Data_collapsars$Prompt.Isotropic.Energy, pch=8, col='blue')
abline(mod1)
#add dashed lines for confidence bands
lines(newx, preds[ ,3], lty = 'dashed', col ='green')
lines(newx, preds[ ,2], lty = 'dashed', col = 'green')
legend("topleft",
       legend = c('Unknown', 'Collapsars'),
       col = c('black', 'blue'),
       pch = c(20, 8))
print("All the work done above is fitting the actual data. Below, I calculate a 99% confidence interval for the slope, (which is our significance level) to see if a linear relationship exists. If zero is in that interval, then the null hypothesis that no such relationship exists cannot be rejected, and I have to move on.")
cat("the upper limit for this interval is: ")
cat(mod1$coefficients[2]-qt(0.99, 48)*mod1$coefficients[1])
cat(" and the lower limit is ")
cat(mod1$coefficients[2]+qt(0.99, 48)*mod1$coefficients[1])
cat(", meaning that the null hypothesis cannot be rejected, and it is possible that no linear relationship exists between the GBM prompt isotropic energy and average XRT x-ray decay rate, despite the result in BAT.")
```

What follows is the plots for the efficiency hypothesis, linking the prompt and afterglow isotropic energies as derived in the python code. As one can see, this relationship does have good significance in addition to its clear physical relevance to both the burst mechanism and the circumburst environment.

```{r Hypothesis 2a, echo=FALSE}
H_2a_Data <- read_excel("/Users/nnuessle/Documents/Work/burst_matching_algorithm/actual_paper/prompt-and-afterglow-matching/Hypothesis_2a_Data.xlsx", col_types = c("skip", "numeric", "numeric", "numeric", "numeric"))
H_2a_Data_copy<-data.frame(H_2a_Data)
# names(H_2a_Data_copy) <- NULL             # Delete column names
# H_2a_Data_copy<-H_2a_Data_copy[-c(15), ] # Print updated data
# cat("we had to remove burst GRB")
# cat(toString(H_2a_Data[15, 'Name']))
# cat(", because it was an outlier in the linear model.")
x<-log10(H_2a_Data_copy[[2]])
y<-log10(H_2a_Data_copy[[3]])
df <- data.frame(x = x, y = y)
mod2a<-lm(y~x)
plot(mod2a)
print(anova(mod2a))
print(summary(mod2a))
#get predicted y values using regression equation
newx <- seq(min(df$x), max(df$x), length.out=100)
preds <- predict(mod2a, newdata = data.frame(x=newx), interval = 'prediction')
par(mar=c(5,5,5,5))
par(mfrow=c(1,1))
collapsar_mask=H_2a_Data_copy$Name %in% Collapsars
merger_mask=H_2a_Data_copy$Name %in% Mergers
H_2a_Data_collapsars=H_2a_Data_copy[collapsar_mask, ]
H_2a_Data_mergers=H_2a_Data_copy[merger_mask, ]
H_2a_Data_unknown=H_2a_Data_copy[!c(collapsar_mask+merger_mask), ]
long_mask=which(H_2a_Data_unknown$t90>4.5)
short_mask=which(H_2a_Data_unknown$t90<=4.5)
H_2a_Data_Long=H_2a_Data_unknown[long_mask, ]
H_2a_Data_Short=H_2a_Data_unknown[short_mask, ]
plot(H_2a_Data_Long$Plateau.E_iso~H_2a_Data_Long$Prompt.E_iso, pch=20, xlab='Prompt Isotropic Energy 10kev-10MeV (erg, GBM)', ylab=paste(strwrap('Plateau Afterglow Isotropic Energy 0.2-10 keV (erg, XRT)', width = 50), collapse = "\n"), log="xy", col='black')
points(H_2a_Data_Short$Plateau.E_iso~H_2a_Data_Short$Prompt.E_iso, pch=20, col='grey')
# points(H_2a_Data_mergers$Plateau.E_iso~H_2a_Data_mergers$Prompt.E_iso, pch=4, col='red', log="xy")
points(H_2a_Data_collapsars$Plateau.E_iso~H_2a_Data_collapsars$Prompt.E_iso, pch=8, col='blue')
abline(mod2a)
#add dashed lines for confidence bands
lines(10^newx, 10^preds[ ,3], lty = 'dashed', col = 'green')
lines(10^newx, 10^preds[ ,2], lty = 'dashed', col = 'green')
legend("topleft",
       legend = c('Unknown Long', 'Unknown Short', 'Collapsars'),
       col = c('black', 'grey', 'blue'),
       pch = c(20, 20, 8))

print("What follows is a version of the model where the lowest-energy prompt emission is removed, to eliminate some of the quartile error. This isn't necessary to bring the model to our predetermined significance level of 1%, nor does it have any physical meaning. Therefore, it is not discussed in the results of the paper.")

new_df <- subset(df, x > 51.5) 
m=new_df$y
n=new_df$x
mod2a1<-lm(m~n)
par(mfrow=c(2,2))
plot(mod2a1)
print(anova(mod2a1))
print(summary(mod2a1))
#get predicted y values using regression equation
newx <- seq(min(n), max(n), length.out=100)
preds <- predict(mod2a1, newdata=data.frame(n=newx), interval='prediction')
par(mar=c(5,5,5,5))
par(mfrow=c(1,1))
H_2a_Data_collapsars_2=subset(H_2a_Data_collapsars, log10(H_2a_Data_collapsars$Prompt.E_iso) > 51.5)
H_2a_Data_unknown_2=subset(H_2a_Data_unknown, log10(H_2a_Data_unknown$Prompt.E_iso) > 51.5)
plot(H_2a_Data_unknown_2$Plateau.E_iso~H_2a_Data_unknown_2$Prompt.E_iso, pch=20, xlab='Prompt Isotropic Energy 10kev-10MeV (erg, GBM)', ylab=paste(strwrap('Plateau Afterglow Isotropic Energy 0.2-10 keV (erg, XRT)', width = 50), collapse = "\n"), log="xy")
# points(H_2a_Data_mergers$Plateau.E_iso~H_2a_Data_mergers$Prompt.E_iso, pch=4, col='red', log="xy")
points(H_2a_Data_collapsars_2$Plateau.E_iso~H_2a_Data_collapsars_2$Prompt.E_iso, pch=8, col='blue')
abline(mod2a1)
#add dashed lines for confidence bands
lines(10^newx, 10^preds[ ,3], lty = 'dashed', col = 'green')
lines(10^newx, 10^preds[ ,2], lty = 'dashed', col = 'green')
legend("topleft",
       legend = c('Unknown', 'Collapsars'),
       col = c('black', 'blue'),
       pch = c(20, 8))

s3d <-scatterplot3d(x=log10(H_2a_Data_unknown$Prompt.E_iso), y=log10(H_2a_Data_unknown$Plateau.E_iso), z=log10(H_2a_Data_unknown$t90), pch=20, color ='black', xlab='Prompt Isotropic Energy 10kev-10MeV (erg, GBM)', ylab=paste(strwrap('Plateau Afterglow Isotropic Energy 0.2-10 keV (erg, XRT)', width = 25), collapse = "\n"), zlab="t90 (s, GBM)") 
s3d$points3d(x=log10(H_2a_Data_collapsars$Prompt.E_iso), y=log10(H_2a_Data_collapsars$Plateau.E_iso), z=log10(H_2a_Data_collapsars$t90), pch=8, col ='blue')
# plot_ly(x=H_2a_Data_copy$Prompt.E_iso, y=H_2a_Data_copy$Plateau.E_iso, z=log10(H_2a_Data_copy$t90), color=log10(H_2a_Data_copy$t90), colors = c('green', 'blue'))
```

```{r Hypothesis 2b, echo=FALSE, warning=FALSE, eval=FALSE}
H_2b_Data <- read_excel("/Users/nnuessle/Documents/Work/burst_matching_algorithm/actual_paper/prompt-and-afterglow-matching/Hypothesis_2b_Data.xlsx", col_types = c("skip", "numeric", "numeric", "numeric", "numeric"))
H_2b_Data_copy<-data.frame(H_2b_Data)
#names(H_2b_Data_copy) <- NULL             # Delete column names
long_list<-which(H_2b_Data_copy[[4]]>4.2)
short_list<-which(H_2b_Data_copy[[4]]<=4.2)
long_H_2b<-H_2b_Data_copy[long_list,]
short_H_2b<-H_2b_Data_copy[short_list,]
collapsar_mask<-long_H_2b$Name %in% Collapsars
merger_mask<-long_H_2b$Name %in% Mergers
H_2b_Data_long_collapsars<-long_H_2b[collapsar_mask, ]
H_2b_Data_long_mergers<-long_H_2b[merger_mask, ]
H_2b_Data_long_unknown<-long_H_2b[!c(collapsar_mask+merger_mask), ]
collapsar_mask<-short_H_2b$Name %in% Collapsars
merger_mask<-short_H_2b$Name %in% Mergers
H_2b_Data_short_collapsars<-short_H_2b[collapsar_mask, ]
H_2b_Data_short_mergers<-short_H_2b[merger_mask, ]
H_2b_Data_short_unknown<-short_H_2b[!c(collapsar_mask+merger_mask), ]
# long_x<-log10(long_H_2b[[2]])
# long_y<-log10(long_H_2b[[3]])
# short_x<-log10(short_H_2b[[2]])
# short_y<-log10(short_H_2b[[3]])
plot(H_2b_Data_long_unknown$End.Plateau.Flux~H_2b_Data_long_unknown$Prompt.Fluence, col='black', pch=20, xlab="Prompt Fluence, 10keV-1MeV (erg cm^-2, GBM)", ylab="End Plateau Flux, 0.2-10 keV (erg s^-1 cm^-2, XRT)", log = "xy", main="The Long Burst Prompt Fluence versus End-Plateau Flux", xlim=c(5e-8, 5e-3))
points(H_2b_Data_long_mergers$End.Plateau.Flux~H_2b_Data_long_mergers$Prompt.Fluence, col='red', pch=4)
points(H_2b_Data_long_collapsars$End.Plateau.Flux~H_2b_Data_long_collapsars$Prompt.Fluence, col='blue', pch=8)
legend("topleft",
       legend = c('Unknown', 'Mergers', 'Collapsars'),
       col = c('black','red', 'blue'),
       pch = c(20, 4, 8))
plot(H_2b_Data_short_unknown$End.Plateau.Flux~H_2b_Data_short_unknown$Prompt.Fluence, col='black', pch=20, xlab="Prompt Fluence, 10keV-1MeV (erg cm^-2, GBM)", ylab="End Plateau Flux, 0.2-10 keV (erg s^-1 cm^-2, XRT)", log = "xy", main="The Short Burst Prompt Fluence versus End-Plateau Flux")
points(H_2b_Data_short_mergers$End.Plateau.Flux~H_2b_Data_short_mergers$Prompt.Fluence, col='red', pch=4)
points(H_2b_Data_short_collapsars$End.Plateau.Flux~H_2b_Data_short_collapsars$Prompt.Fluence, col='blue', pch=8)
legend("topleft",
       legend = c('Unknown'),
       col = c('black'),
       pch = c(20, 4, 8))
plot(c(H_2b_Data_long_unknown$End.Plateau.Flux,H_2b_Data_short_unknown$End.Plateau.Flux)~c(H_2b_Data_long_unknown$Prompt.Fluence,H_2b_Data_short_unknown$Prompt.Fluence), col='black', pch=20, xlab="Prompt Fluence, 10keV-1MeV (erg cm^-2, GBM)", ylab="End Plateau Flux, 0.2-10 keV (erg s^-1 cm^-2, XRT)", log = "xy", main="All Burst Prompt Fluences versus End-Plateau Fluxes", xlim=c(5e-8, 5e-3))
points(c(H_2b_Data_long_mergers$End.Plateau.Flux,H_2b_Data_short_mergers$End.Plateau.Flux)~c(H_2b_Data_long_mergers$Prompt.Fluence,H_2b_Data_short_mergers$Prompt.Fluence), col='red', pch=4)
points(c(H_2b_Data_long_collapsars$End.Plateau.Flux,H_2b_Data_short_collapsars$End.Plateau.Flux)~c(H_2b_Data_long_collapsars$Prompt.Fluence,H_2b_Data_short_collapsars$Prompt.Fluence), col='blue', pch=8)
legend("topleft",
       legend = c('Unknown', 'Mergers', 'Collapsars'),
       col = c('black','red', 'blue'),
       pch = c(20, 4, 8))
s3d <-scatterplot3d(x=log10(H_2b_Data_long_unknown$End.Plateau.Flux), y=log10(H_2b_Data_long_unknown$End.Plateau.Flux), z=log10(H_2b_Data_long_unknown$t90), color="black", pch=20, xlab="Prompt Fluence, 10keV-1MeV (erg cm^-2, GBM)", ylab=paste(strwrap("End Plateau Flux, 0.2-10 keV (erg s^-1 cm^-2, XRT)", width = 25), collapse = "\n"), zlab="t90 (s, GBM)", main = paste(strwrap("Three-Dimensional Plot of all Data points Motivating Machine Learning by Separation of Long and Short Bursts, Though Not Progenitors", width = 50), collapse = "\n"), xlim = c(-14, -3), zlim = c(-1, 3))
s3d$points3d(x=log10(H_2b_Data_short_unknown$Prompt.Fluence), y=log10(H_2b_Data_short_unknown$End.Plateau.Flux), z=log10(H_2b_Data_short_unknown$t90), col="grey", pch=20)
s3d$points3d(x=c(log10(H_2b_Data_long_mergers$Prompt.Fluence),log10(H_2b_Data_short_mergers$Prompt.Fluence)), y=c(log10(H_2b_Data_long_mergers$End.Plateau.Flux),log10(H_2b_Data_short_mergers$End.Plateau.Flux)), z=c(log10(H_2b_Data_long_mergers$t90),log10(H_2b_Data_short_mergers$t90)), col="red", pch=4)
s3d$points3d(x=c(log10(H_2b_Data_long_collapsars$Prompt.Fluence),log10(H_2b_Data_short_collapsars$Prompt.Fluence)), y=c(log10(H_2b_Data_long_collapsars$End.Plateau.Flux),log10(H_2b_Data_short_collapsars$End.Plateau.Flux)), z=c(log10(H_2b_Data_long_collapsars$t90),log10(H_2b_Data_short_collapsars$t90)), col="blue", pch=8)
legend("topleft",
       legend = c('Unknown (Long)', 'Unknown (Short)', 'Mergers', 'Collapsars'),
       col = c('black','grey','red', 'blue'),
       pch = c(20,20, 4, 8))
missing<-which(!complete.cases(H_2b_Data_copy))
no_NA_H_2b_Data<- H_2b_Data_copy[-missing, ]
collapsar_mask<-no_NA_H_2b_Data$Name %in% Collapsars
merger_mask<-no_NA_H_2b_Data$Name %in% Mergers
short_list<-which(no_NA_H_2b_Data$t90<4.2)
long_list<-which(no_NA_H_2b_Data$t90>4.2)
Classification<-rep("Unknown", nrow(no_NA_H_2b_Data))
Classification[short_list]<-"merger"
Classification[long_list]<-"collapsar" 
Classification[collapsar_mask]<-"collapsar" #It kinda looks like an explosion I guess IDK
Classification[merger_mask]<-"merger" #and this can be the jet of the merger
new_H_2b_Data<-data.frame(cbind(log10(no_NA_H_2b_Data$Prompt.Fluence), log10(no_NA_H_2b_Data$End.Plateau.Flux), log10(no_NA_H_2b_Data$t90), as.factor(Classification)))
colnames(new_H_2b_Data)<-c("Prompt.Fluence", "End.Plateau.Flux", "t90", "as.factor.Classification.")
# model <- randomForest(formula = no_NA_H_2b_Data$Classification ~ ., data = no_NA_H_2b_Data)
svmfit<-svm(new_H_2b_Data$as.factor.Classification. ~ ., data = new_H_2b_Data, type = "C-classification", cost = 1.0, kernel = "linear")
print(svmfit)
plot(svmfit, new_H_2b_Data, End.Plateau.Flux~Prompt.Fluence, slice = list(t90 = 0.5, t90 = 1), xlab="Prompt Fluence, 10keV-1MeV (erg cm^-2, GBM)", ylab=paste(strwrap("End Plateau Flux, 0.2-10 keV (erg s^-1 cm^-2, XRT)", width = 25), collapse = "\n"))
plot(svmfit, new_H_2b_Data, End.Plateau.Flux~t90, slice = list(Prompt.Fluence = -5.5, Prompt.Fluence = -4.5), xlab="t90 (s, GBM)", ylab=paste(strwrap("End Plateau Flux, 0.2-10 keV (erg s^-1 cm^-2, XRT)", width = 25), collapse = "\n"))
plot(svmfit, new_H_2b_Data, Prompt.Fluence~t90, slice = list(End.Plateau.Flux = -10, End.Plateau.Flux  = -9), xlab="Prompt Fluence, 10keV-1MeV (erg cm^-2, GBM)",  ylab="t90 (s, GBM)",)
# Simulate some data
n    = 100
nnew = 50
set.seed(12345)
group = sample(2, n, replace=T)
dat   = data.frame(group=factor(group), matrix(rnorm(n*3, rep(group, each=3)), ncol=3, byrow=T))

# # Fit SVM
# fit = svm(group ~ ., data=dat)
# 
# # Plot original data
# plot3d(dat[,-1], col=dat$group)

# Get decision values for a new data grid
newdat.list = lapply(new_H_2b_Data, function(x) seq(min(x), max(x), len=nnew))
newdat      = expand.grid(newdat.list)
names(newdat) <- c("Prompt.Fluence", "End.Plateau.Flux", "t90")
newdat.pred = predict(svmfit, newdata=newdat, decision.values=T)
newdat.dv   = attr(newdat.pred, 'decision.values')
newdat.dv   = array(newdat.dv, dim=rep(nnew, 3))

# Fit/plot an isosurface to the decision boundary
plot3d(x=new_H_2b_Data$Prompt.Fluence, y=new_H_2b_Data$End.Plateau.Flux, z=new_H_2b_Data$t90, xlab="Prompt Fluence, 10keV-1MeV (erg cm^-2, GBM)", ylab=paste(strwrap("End Plateau Flux, 0.2-10 keV (erg s^-1 cm^-2, XRT)", width = 25), collapse = "\n"), zlab="t90 (s, GBM)")
contour3d(newdat.dv, level=0, x=newdat.list$Prompt.Fluence, y=newdat.list$End.Plateau.Flux, z=newdat.list$t90, add=T)
```

```{r Goldstein Hypothesis, echo=FALSE}
Goldstein_S2_Data <- read_excel("/Users/nnuessle/Documents/Work/burst_matching_algorithm/actual_paper/prompt-and-afterglow-matching/Goldstein_Sample_2_Data.xlsx", col_types = c("skip", "numeric", "numeric", "numeric", "numeric")) [-c(3), ]
par(mar=c(5,5,5,5))
collapsar_mask=Goldstein_S2_Data$Name %in% Collapsars
merger_mask=Goldstein_S2_Data$Name %in% Mergers
Goldstein_S2_collapsars=Goldstein_S2_Data[collapsar_mask, ]
Goldstein_S2_mergers=Goldstein_S2_Data[merger_mask, ]
Goldstein_S2_unknown=Goldstein_S2_Data[!c(collapsar_mask+merger_mask), ]
long_mask=which(Goldstein_S2_unknown$t90>4.5)
short_mask=which(Goldstein_S2_unknown$t90<=4.5)
Goldstein_S2_Long=Goldstein_S2_unknown[long_mask, ]
Goldstein_S2_Short=Goldstein_S2_unknown[short_mask, ]
plot(Goldstein_S2_Long$t90, Goldstein_S2_Long$`Peak E. over Flue.`, pch=20, ylab=expression('E_peak/Fluence 10kev-10MeV ('~erg ~ cm^{2} ~ erg^{-1} ~ ', GBM)'), xlab='t90 (s, GBM)', log="xy", xlim=c(1e-2, 1e3))
points(Goldstein_S2_Short$t90, Goldstein_S2_Short$`Peak E. over Flue.`, pch=20, col='grey')
points(Goldstein_S2_mergers$t90, Goldstein_S2_mergers$`Peak E. over Flue.`, pch=4, col='red')
points(Goldstein_S2_collapsars$t90, Goldstein_S2_collapsars$`Peak E. over Flue.`, pch=8, col='blue')
legend("topleft",
       legend = c('Unknown Long', 'Unknown Short', "Mergers", 'Collapsars'),
       col = c('black', 'grey', 'red', 'blue'),
       pch = c(20, 20, 4, 8))
##
missing<-which(!complete.cases(Goldstein_S2_Data))
no_NA_Goldstein_S2_Data<- Goldstein_S2_Data[-missing, ]
collapsar_mask<-no_NA_Goldstein_S2_Data$Name %in% Collapsars
merger_mask<-no_NA_Goldstein_S2_Data$Name %in% Mergers
# short_list<-which(no_NA_Goldstein_S2_Data$t90<4.2)
# long_list<-which(no_NA_Goldstein_S2_Data$t90>4.2)
Classification<-rep("Unknown", nrow(no_NA_Goldstein_S2_Data))
# Classification[short_list]<-"merger"
# Classification[long_list]<-"collapsar" 
Classification[collapsar_mask]<-"collapsar"
Classification[merger_mask]<-"merger"
new_Goldstein_S2_Data<-data.frame(cbind(log10(no_NA_Goldstein_S2_Data$`Peak E. over Flue.`), log10(no_NA_Goldstein_S2_Data$t90), as.factor(Classification)))
colnames(new_Goldstein_S2_Data)<-c("Peak_E_over_Flu", "t90", "as.factor.Classification.")
# model <- randomForest(formula = no_NA_H_2b_Data$Classification ~ ., data = no_NA_H_2b_Data)
unknown_remover<-which(new_Goldstein_S2_Data$as.factor.Classification.==3)
merger_mask<-which(new_Goldstein_S2_Data$as.factor.Classification.==2)
collapsar_mask<-which(new_Goldstein_S2_Data$as.factor.Classification.==1)
training_set=new_Goldstein_S2_Data[-unknown_remover, ]
svmfit<-svm(training_set$as.factor.Classification. ~ ., data = training_set, type = "C-classification", cost = 1.0, kernel = "sigmoid")
print(svmfit)
plot(svmfit, training_set, Peak_E_over_Flu~t90, ylab=expression('E_peak/Fluence 10kev-10MeV ('~erg ~ cm^{2} ~ erg^{-1} ~ ', GBM)'), xlab='t90 (s, GBM)', log="xy")
# points(log10(Goldstein_S2_unknown$t90), log10(Goldstein_S2_unknown$`Peak E. over Flue.`))
predicted_svm<-predict(svmfit, new_Goldstein_S2_Data[unknown_remover, ])
plot(new_Goldstein_S2_Data[unknown_remover, ]$Peak_E_over_Flu~new_Goldstein_S2_Data[unknown_remover, ]$t90, ylab=expression('E_peak/Fluence 10kev-10MeV ('~erg ~ cm^{2} ~ erg^{-1} ~ ', GBM)'), xlab='t90 (s, GBM)', col=predicted_svm, pch=20)
points(log10(Goldstein_S2_mergers$t90), log10(Goldstein_S2_mergers$`Peak E. over Flue.`), pch=4, col='#09622A')
points(log10(Goldstein_S2_collapsars$t90), log10(Goldstein_S2_collapsars$`Peak E. over Flue.`), pch=8, col='blue')
legend("topleft",
       legend = c('Sorted Collapsars', 'Sorted Mergers', "Detected Mergers", 'Detected Collapsars'),
       col = c('black', 'red', '#09622A', 'blue'),
       pch = c(20, 20, 4, 8))
###
#now let's try it on the overlapping sample with redshifts
Goldstein_S6_Data <- read_excel("/Users/nnuessle/Documents/Work/burst_matching_algorithm/actual_paper/prompt-and-afterglow-matching/Goldstein_Sample_6_Data.xlsx", col_types = c("skip", "numeric", "numeric", "numeric", "numeric"))
missing<-which(!complete.cases(Goldstein_S6_Data))
no_NA_Goldstein_S6_Data<- Goldstein_S6_Data[-missing, ]
collapsar_mask<-no_NA_Goldstein_S6_Data$Name %in% Collapsars
merger_mask<-no_NA_Goldstein_S6_Data$Name %in% Mergers
Classification<-rep("Unknown", nrow(no_NA_Goldstein_S6_Data))
Classification[collapsar_mask]<-"collapsar"
Classification[merger_mask]<-"merger"
new_Goldstein_S6_Data<-data.frame(cbind(log10(no_NA_Goldstein_S6_Data$`Peak E. over Flue.`), log10(no_NA_Goldstein_S6_Data$t90), as.factor(Classification)))
colnames(new_Goldstein_S6_Data)<-c("Peak_E_over_Flu", "t90", "as.factor.Classification.")
unknown_remover<-which(new_Goldstein_S6_Data$as.factor.Classification.==3)
Second_predicted_svm<-predict(svmfit, new_Goldstein_S6_Data)
plot(new_Goldstein_S6_Data$Peak_E_over_Flu~new_Goldstein_S6_Data$t90, ylab=expression('E_peak/Fluence 10kev-10MeV ('~erg ~ cm^{2} ~ erg^{-1} ~ ', GBM)'), xlab='t90 (s, GBM)', col=Second_predicted_svm, pch=20)
svm_check=predict(svmfit, new_Goldstein_S6_Data[-unknown_remover, ])
confusionMatrix(svm_check, factor(new_Goldstein_S6_Data[-unknown_remover, ]$as.factor.Classification.), dnn = c("Prediction", "Reference"))
```

```{r Gehrels Hypothesis, echo=FALSE}
Gehrels_S3_Data <- read_excel("/Users/nnuessle/Documents/Work/burst_matching_algorithm/actual_paper/prompt-and-afterglow-matching/Gehrels_Sample_3_Data.xlsx", col_types = c("skip", "numeric", "numeric", "numeric", "numeric"))
Gehrels_S3_Data<-Gehrels_S3_Data[-c(146), ] #removed an outlier from the linear model
collapsar_mask=Gehrels_S3_Data$Name %in% Collapsars
merger_mask=Gehrels_S3_Data$Name %in% Mergers
Gehrels_S3_collapsars=Gehrels_S3_Data[collapsar_mask, ]
Gehrels_S3_mergers=Gehrels_S3_Data[merger_mask, ]
Gehrels_S3_unknown=Gehrels_S3_Data[!c(collapsar_mask+merger_mask), ]
long_mask=which(Gehrels_S3_unknown$t90>4.5)
short_mask=which(Gehrels_S3_unknown$t90<=4.5)
Gehrels_S3_Long=Gehrels_S3_unknown[long_mask, ]
Gehrels_S3_Short=Gehrels_S3_unknown[short_mask, ]
x<-log10(Gehrels_S3_Data$Prompt_Flu)
y<-log10(Gehrels_S3_Data$`X-ray_F11`)
z<-log10(Gehrels_S3_Data$t90)
x2<-log10(Gehrels_S3_Long$Prompt_Flu)
y2<-log10(Gehrels_S3_Long$`X-ray_F11`)
z2<-log10(Gehrels_S3_Long$t90)
##
modGehrels<-lm(y~x)
plot(modGehrels)
print(anova(modGehrels))
print(summary(modGehrels))
#
modGehrels2<-lm(y~x+z)
par(mfrow=c(2,2))
plot(modGehrels2)
print(anova(modGehrels2))
print(summary(modGehrels2))
#
modGehrels3<-lm(y2~x2)
par(mfrow=c(2,2))
plot(modGehrels3)
print(anova(modGehrels3))
print(summary(modGehrels3))
##
newx <- seq(min(x), max(x), length.out=100)
preds <- predict(modGehrels, newdata=data.frame(x=newx), interval='prediction')
##
par(mfrow=c(1,1))
par(mar=c(5,5,5,5))
plot(Gehrels_S3_unknown$Prompt_Flu, Gehrels_S3_unknown$`X-ray_F11`, pch=20, xlab=expression('Prompt Fluence 10kev-10MeV (erg'~cm^{-2}~', GBM)'), ylab=expression('Afterglow '~Flux[11~Hr.]~' 0.2-10 keV (erg'~cm^{-2}~s^{-1}~', XRT)'), log="xy", xlim = c(1e-7, 1e-3), ylim = c(1e-20, 1e-8))
points(Gehrels_S3_mergers$Prompt_Flu, Gehrels_S3_mergers$`X-ray_F11`, pch=4, col='red')
points(Gehrels_S3_collapsars$Prompt_Flu, Gehrels_S3_collapsars$`X-ray_F11`, pch=8, col='blue')
abline(modGehrels)
#add dashed lines for confidence bands
lines(10^newx, 10^preds[ ,3], lty = 'dashed', col = 'green')
lines(10^newx, 10^preds[ ,2], lty = 'dashed', col = 'green')
legend("bottomleft",
       legend = c('Unknown', 'Mergers', 'Collapsars'),
       col = c('black', 'red', 'blue'),
       pch = c(20, 4, 8))
##
# newx <- seq(min(x2), max(x2), length.out=100)
# preds <- predict(modGehrels, newdata=data.frame(x2=newx), interval='prediction')
##
par(mar=c(5,5,5,5))
plot(Gehrels_S3_Long$Prompt_Flu, Gehrels_S3_Long$`X-ray_F11`, pch=20, col='black', xlab=expression('Prompt Fluence 10kev-10MeV (erg'~cm^{-2}~', GBM)'), ylab=expression('Afterglow '~Flux[11~Hr.]~' 0.2-10 keV (erg'~cm^{-2}~s^{-1}~', XRT)'), log="xy", xlim = c(1e-7, 1e-3), ylim = c(1e-20, 1e-8))
points(Gehrels_S3_Short$Prompt_Flu, Gehrels_S3_Short$`X-ray_F11`, pch=20, col='grey')
points(Gehrels_S3_mergers$Prompt_Flu, Gehrels_S3_mergers$`X-ray_F11`, pch=4, col='red')
points(Gehrels_S3_collapsars$Prompt_Flu, Gehrels_S3_collapsars$`X-ray_F11`, pch=8, col='blue')
abline(modGehrels)
#add dashed lines for confidence bands
lines(10^newx, 10^preds[ ,3], lty = 'dashed', col = 'green')
lines(10^newx, 10^preds[ ,2], lty = 'dashed', col = 'green')
legend("bottomleft",
       legend = c('Unknown Long', "Unknown Short", 'Mergers', 'Collapsars'),
       col = c('black', 'grey', 'red', 'blue'),
       pch = c(20, 20, 4, 8))
####
Gehrels_S6_Data <- read_excel("/Users/nnuessle/Documents/Work/burst_matching_algorithm/actual_paper/prompt-and-afterglow-matching/Gehrels_Sample_6_Data.xlsx", col_types = c("skip", "numeric", "numeric", "numeric", "numeric"))
Gehrels_S6_Data<-Gehrels_S6_Data[-c(40), ] #removed an outlier from the linear model
collapsar_mask=Gehrels_S6_Data$Name %in% Collapsars
merger_mask=Gehrels_S6_Data$Name %in% Mergers
Gehrels_S6_collapsars=Gehrels_S6_Data[collapsar_mask, ]
Gehrels_S6_mergers=Gehrels_S6_Data[merger_mask, ]
Gehrels_S6_unknown=Gehrels_S6_Data[!c(collapsar_mask+merger_mask), ]
long_mask=which(Gehrels_S6_unknown$t90>4.5)
short_mask=which(Gehrels_S6_unknown$t90<=4.5)
Gehrels_S6_Long=Gehrels_S6_unknown[long_mask, ]
Gehrels_S6_Short=Gehrels_S6_unknown[short_mask, ]
x<-log10(Gehrels_S6_Data$Prompt_E_iso)
y<-log10(Gehrels_S6_Data$`X-ray_F11`)
z<-log10(Gehrels_S6_Data$t90)
x2<-log10(Gehrels_S6_Long$Prompt_E_iso)
y2<-log10(Gehrels_S6_Long$`X-ray_F11`)
z2<-log10(Gehrels_S6_Long$t90)
##
modGehrels<-lm(y~x)
par(mfrow=c(2,2))
plot(modGehrels)
print(anova(modGehrels))
print(summary(modGehrels))
#
modGehrels2<-lm(y~x+z)
plot(modGehrels2)
print(anova(modGehrels2))
print(summary(modGehrels2))
#
modGehrels3<-lm(y2~x2)
par(mfrow=c(2,2))
plot(modGehrels3)
print(anova(modGehrels3))
print(summary(modGehrels3))
##
newx <- seq(min(x), max(x), length.out=100)
preds <- predict(modGehrels, newdata=data.frame(x=newx), interval='prediction')
##
par(mfrow=c(1,1))
par(mar=c(5,5,5,5))
plot(Gehrels_S6_unknown$Prompt_E_iso, Gehrels_S6_unknown$`X-ray_F11`, pch=20,  xlab=expression('Prompt E_iso 10kev-10MeV (erg, GBM)'), ylab=expression('Afterglow '~Flux[11~Hr.]~' 0.2-10 keV (erg'~cm^{-2}~s^{-1}~', XRT)'), log="xy", ylim = c(1e-16, 1e-8))
# points(Gehrels_S6_mergers$Prompt_E_iso, Gehrels_S6_mergers$`X-ray_F11`, pch=4, col='red')
points(Gehrels_S6_collapsars$Prompt_E_iso, Gehrels_S6_collapsars$`X-ray_F11`, pch=8, col='blue')
abline(modGehrels)
#add dashed lines for confidence bands
lines(10^newx, 10^preds[ ,3], lty = 'dashed', col = 'green')
lines(10^newx, 10^preds[ ,2], lty = 'dashed', col = 'green')
legend("bottomleft",
       legend = c('Unknown', 'Collapsars'),
       col = c('black', 'blue'),
       pch = c(20, 8))
##
# newx <- seq(min(x2), max(x2), length.out=100)
# preds <- predict(modGehrels, newdata=data.frame(x2=newx), interval='prediction')
##
par(mar=c(5,5,5,5))
plot(Gehrels_S6_Long$Prompt_E_iso, Gehrels_S6_Long$`X-ray_F11`, pch=20, col='black', xlab=expression('Prompt E_iso 10kev-10MeV (erg, GBM)'), ylab=expression('Afterglow '~Flux[11~Hr.]~' 0.2-10 keV (erg'~cm^{-2}~s^{-1}~', XRT)'), log="xy", ylim = c(1e-16, 1e-8))
points(Gehrels_S6_Short$Prompt_E_iso, Gehrels_S6_Short$`X-ray_F11`, pch=20, col='grey')
# points(Gehrels_S3_mergers$Prompt_Flu, Gehrels_S3_mergers$`X-ray_F11`, pch=4, col='red')
points(Gehrels_S6_collapsars$Prompt_E_iso, Gehrels_S6_collapsars$`X-ray_F11`, pch=8, col='blue')
abline(modGehrels)
#add dashed lines for confidence bands
lines(10^newx, 10^preds[ ,3], lty = 'dashed', col = 'green')
lines(10^newx, 10^preds[ ,2], lty = 'dashed', col = 'green')
legend("bottomleft",
       legend = c('Unknown Long', "Unknown Short", 'Collapsars'),
       col = c('black', 'grey', 'blue'),
       pch = c(20, 20, 8))
```

